Model Adaptivity for Goal-Oriented Inference using Adjoints

### 2017-06-07 paper/submitted/round2_png/reviewer_response.tex

\newcommand{\answer}[1]{\begin{adjustwidth}{1cm}{}{\color{blue}#1}\end{adjustwidth}}
\newcommand{\red}[1]{{\color{red}#1}}
\newcommand{\blue}[1]{{\color{blue}#1}}
\newcommand{\green}[1]{{\color{green}#1}}

\newcommand{\notdone}{{\color{red}{Changes not yet made.}}}
\newcommand{\done}{{\color{green}{Changes made.}}}

\begin{document}

%%%%% ASK KAREN/ELIZABETH
%do we need to refer to line numbers?
%if a comment asks for clarification, do we reproduce it here or point to the changed lines/sections?

@@

I found the motivation in the beginning of the manuscript somewhat lacking. The description of use
cases, and conditions for which the algorithm is suitable, from the last paragraph on page 9 were very
helpful. I believe the presentation would be strengthened by mentioning some of this earlier in the
introduction.

\answer{We added a sentence about the use cases to the second paragraph of the introduction.}

Although it is represented graphically in Figure 8, $\Omega_I$ for the final example is not defined in the text.

\answer{It is now defined just before Figure 8 (now Figure 9) is referenced.}

@@

At the beginning of the second paragraph of Section 4.3.2 (pg. 25) I think the reference to “Section
4.3.2” should be Table 1.

\answer{Thanks for the catch. We've corrected the reference.}

The sub-figures of Figure 9 could be larger. It is difficult to verify the claim in the text that “. . . the
domain is refined completely in the $x_3$ direction first around the QoI region. . . ”. Even going as far as
to include 2-D slices (as in Figure 8) may be warranted.

\answer{We made the figure (now Figure 10) larger, made the wording of that claim more specific, and included a 2D slice for each mixed-fidelity model that cuts through the QoI region, to aid in visual verification.}

@@

\answer{}
\notdone

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection*{Reviewer 3}

I'd like to actually see the statement and proof of Proposition 3 from reference 7 included in
the manuscript. It is at the crux of the proof and it's frustrating to have to go dig up that
detail.

@@

\answer{}
\notdone

I'd like to see the details of the error estimate localization included in the manuscript instead
of delegating to a reference. The given description is quite “wordy” and not very precise.

\answer{We've added the relevant parts of the reference to Section 3.2.}

The spirit of the examples is fine, but I'm not following some of the details. They state for
the example in Section 4.1 that $f (q) = q$, but then the exact underlying field varies in space.
To what space does $f (q)$ belong? Based on the example in Section 4.2.1, it looks like it's
a constant. If the parameters are spatially dependent, what underlying basis is being used
for its representation? Is the mesh refined enough to rule out pollution of the results due to
discretization error?

\answer{We reworded the description of the parameters in Section 4.2.1 to be more clear/explicit about the relevant spaces and representations. \red{Not sure what to make of discretization error comment...}}
\notdone

@@

It could be illustrative to see how the parameters q change with the refinement process.

\answer{We added an additional figure (now Figure 7) showing the parameters $q$ for some of the refinement iterations, along with commentary.}

### 2017-06-07 paper/submitted/round2_png/mathematical_formulation_v4_cmame.tex
Added distinction between in low and high fidelity models clearly, with separate statements for the two.

\section{Mathematical Formulation}\label{sec:form}
%
We first introduce the goal-oriented inverse problem. Then we derive a rigorous a posteriori error estimate for the error induced in the QoI due to the use of mixed-fidelity and/or lower-fidelity models.

%------------------------------------------------------------------------------%
\subsection{Inverse Problem Formulation}  \label{sec:setup}
%------------------------------------------------------------------------------%
%
Consider a model for which the Galerkin formulation of the weak form is written as
\begin{equation}
a(u,q)(\phi)=\ell(q)(\phi),\quad\forall\phi\in U,
\label{eq:weakForm}
\end{equation}
where $u\in U$ is the state, $q\in Q$ are the unknown parameters, $\phi$ is a test function, and the state space $U$ and parameter space $Q$ are Hilbert spaces. The form $a$ and functional $\ell$ are linear with respect to the arguments in the second pair of parentheses (in \Cref{eq:weakForm}, they are linear with respect to $\phi$). Further, we define an observation operator $C:U\to\Reals^{n_d}$ that maps the state to $n_d$ predicted observations. The actual observations (data) are denoted by $d\in\R^{n_d}$.

The unknown parameter $q$ can be inferred by minimizing the difference between the predicted and actual observations, leading to an inverse problem. In our setting, such inverse problems are typically ill-posed, since the observations are sparse, and insufficiently informative to uniquely determine the parameters. To make the inverse problem well-posed, a regularization, denoted by $R(q)$, is used to inject prior information or beliefs about the parameters into the formulation. The regularized inverse problem can thus be written as a constrained optimization problem,
%
\begin{subequations}
\label{eq:invOpt}
\begin{align}
\min\limits_{q,u} & \quad J(q,u)=\frac{1}{2}\|d-C(u)\|_2^2 + R(q), \label{eq:invOpt_obj} \\
\textrm{s.t. }& \quad a(u,q)(\phi)=\ell(q)(\phi),\quad\forall\phi\in U. \label{eq:invOpt_cons}
\end{align}
\end{subequations}
%
Thus, we aim to minimize the cost function $J$, which includes the mismatch between predicted and actual observations and a regularization term $R(q)$, subject to the state $u$ and parameters $q$ satisfying the model given by \Cref{eq:weakForm}, which appears as a constraint in \Cref{eq:invOpt_cons}. The constraints \Cref{eq:invOpt_cons} are typically models of physical processes or systems. A given physical system can be described to varying degrees of fidelity using different models. This naturally introduces a hierarchy of models and a tradeoff between fidelity and computational expense; how one manages this tradeoff depends on what aspect of the solution is of interest.

We consider the case of a {\em goal-oriented inverse problem}, where the ultimate purpose of inferring the unknown parameters is to calculate some quantity of interest (QoI). We denote a scalar QoI by $I(q,u)$, where $I:Q\times U\to\R$ is a functional that maps the parameters and state to our QoI. We then consider the tradeoff between the error in the QoI and the fidelity (and corresponding computational expense) of the model we use. In the following, we introduce the QoI functional $I$ into the inverse problem formulation by introducing auxiliary variables and additional adjoint equations. We then use this formulation to derive an a posteriori error estimate for $I$, where the errors considered are those due to the use of different multi-fidelity models in the constraint \Cref{eq:invOpt_cons}.
%
%------------------------------------------------------------------------------%
\subsection[Error Estimate for a Goal-Oriented Inverse Problem]{Error Estimate for a Goal-Oriented Inverse Problem}  \label{sec:deriv}
%------------------------------------------------------------------------------%
%
For a given hierarchy of models, consider the QoI calculated from inferring the parameters with the highest-fidelity model; we take this QoI to be the value with which we compare other QoI estimates. In this section we derive an a posteriori estimate for the error in the QoI from inferring the parameters with a lower-fidelity model.
%
\begin{proposition}
\label{thm:error_estimate}
Consider the inverse problem described by the constrained optimization problem \Cref{eq:invOpt}. Let the form $a:U \times U \to \Reals$ be three times continuously differentiable with respect to the state $u$ and parameters $q$. Let the observation operator $C:U\to\Reals^{n_d}$ be three times continuously differentiable with respect to the state $u$. Also, let the regularization operator $R:Q\to\Reals$ be differentiable with respect to the parameter $q$, and the functional $I:Q\times U\to\Reals$ be differentiable with respect to the state $u$ and parameter $q$.

>>>>

\section{Mathematical Formulation}\label{sec:form}
%
\edit{In this section, we clearly define the low, mixed and high fidelity inverse problems. We also develop theoretical results which extend the work of~\cite{} to multimodel settings, in particular obtaining a computable, rigorous error estimate for QoI targeted generation of mixed fidelity formulations of inverse problems.}

%------------------------------------------------------------------------------%
\subsection{Inverse Problem Formulation}  \label{sec:setup}
%------------------------------------------------------------------------------%
%
We begin by defining a generic goal-oriented inverse problem. Given observations (data) $d\in\R^{n_d}$, we seek to infer parameter(s) $q\in Q$, where $Q$ is a Hilbert space. An observation operator $C:U\to\Reals^{n_d}$ relates the parameter to the observations via state variable(s) $u\in U$, where $u$ satisfies,
\begin{equation}
a(u,q)(\phi)=\ell(q)(\phi),\quad\forall\phi\in U,
\label{eq:weakForm}
\end{equation}
where $U$ is also a Hilbert space. Equation \Cref{eq:weakForm} is called the state equation. The form $a$, and the functional $\ell$ are linear with respect to the arguments in the second pair of parentheses (in \Cref{eq:weakForm}, they are linear with respect to $\phi$). Finally, our objective functional or the QoI, is a scalar functional $I(q,u)$, where $I:Q\times U\to\R$.

The unknown parameter $q$ can be inferred by minimizing the difference between the predicted and actual observations, leading to an inverse problem. Such inverse problems are typically ill-posed, since the observations are sparse, and insufficiently informative to uniquely determine the parameters. To make the inverse problem well-posed, a regularization, denoted by $R(q)$, is used to inject prior information or beliefs about the parameters into the formulation. The regularized inverse problem can thus be written as a constrained optimization problem,
%
\begin{subequations}
\label{eq:invOpt}
\begin{align}
\min\limits_{q,u} & \quad J(q,u)=\frac{1}{2}\|d-C(u)\|_2^2 + R(q), \label{eq:invOpt_obj} \\
\textrm{s.t. }& \quad a(u,q)(\phi)=\ell(q)(\phi),\quad\forall\phi\in U. \label{eq:invOpt_cons}
\end{align}
\end{subequations}
%
Thus, we aim to minimize the cost function $J$, which includes the mismatch between predicted and actual observations and a regularization term $R(q)$, subject to the state $u$ and parameters $q$ satisfying the model given by \Cref{eq:weakForm}, which appears as a constraint in \Cref{eq:invOpt_cons}.

In an inverse problem, multiple choices are apparent for the structure of the inferred parameter $q$, the data $d$ and state equation. Together these constitute a model for formulating the inverse problem. For example, the parameter $q$ can be modeled as a single scalar value or an infinite dimensional field variable, the data $d$ can be the entire time series of measurements from a sensor, or simply a smaller subset of those measurements. The state equation can be complex, nonlinear relationship requiring an expensive computation to resolve (for e.g./ the Navier Stokes equations) or a linear, single physics diffusion model (for e.g./ Laplace's equation). All these different modeling choices lead to different versions of the optimization problem that would need to be solved to infer the unknown parameter.

We assume that there exists an ideal model for solving the inverse problem, called the high fidelity model. Giving infinite computational resources, we would use this high fidelity to formulate and solve our inverse problem. Mathematically we differentiate this high fidelity inverse problem with the subscript $HF$,
%
\begin{subequations}
\label{eq:invOpt}
\begin{align}
\min\limits_{q,u} & \quad J(q_{HF},u_{HF})=\frac{1}{2}\|d-C(u_{HF})\|_2^2 + R(q_{HF}), \label{eq:invOpt_obj} \\
\textrm{s.t. }& \quad a(u_{HF},q_{HF})(\phi)=\ell(q_{HF})(\phi),\quad\forall\phi\in U_{HF}. \label{eq:invOpt_cons}
\end{align}
\end{subequations}
%
where $q \in Q_{HF}$. Due to the constraints presented by limited computational resources, one may instead attempt to infer the parameter $q$ using a different, less computationally demanding model, which we call the low fidelity model,
%
\begin{subequations}
\label{eq:invOpt}
\begin{align}
\min\limits_{q,u} & \quad J(q_{LF},u_{LF})=\frac{1}{2}\|d-C(u_{LF})\|_2^2 + R(q_{LF}), \label{eq:invOpt_obj} \\
\textrm{s.t. }& \quad a(u_{LF},q_{LF})(\phi)=\ell(q_{LF})(\phi),\quad\forall\phi\in U_{LF}. \label{eq:invOpt_cons}
\end{align}
\end{subequations}
%
where $q \in Q_{LF}$. If there is compatibility between the low fidelity and high fidelity solution spaces for $q$ and $u$, one may combine the two models to solve the inverse problem. The low fidelity model can be used in part of the computational domain, and the high fidelity in the remainder, and such a model is called a mixed fidelity model, which we distinguish by the subscript $MF$.

We further consider the case of a {\em goal-oriented inverse problem}, where the ultimate purpose of inferring the unknown parameters is to calculate some quantity of interest (QoI). We denote a scalar QoI by  is a functional that maps the parameters and state to our QoI. We then consider the tradeoff between the error in the QoI and the fidelity (and corresponding computational expense) of the model we use. In the following, we introduce the QoI functional $I$ into the inverse problem formulation by introducing auxiliary variables and additional adjoint equations. We then use this formulation to derive an a posteriori error estimate for $I$, where the errors considered are those due to the use of different multi-fidelity models in the constraint \Cref{eq:invOpt_cons}.
%
%------------------------------------------------------------------------------%
\subsection[Error Estimate for a Goal-Oriented Inverse Problem]{Error Estimate for a Goal-Oriented Inverse Problem}  \label{sec:deriv}
%------------------------------------------------------------------------------%
%
For a given hierarchy of models, consider the QoI calculated from inferring the parameters with the highest-fidelity model; we take this QoI to be the value with which we compare other QoI estimates. In this section we derive an a posteriori estimate for the error in the QoI from inferring the parameters with a lower-fidelity model.
%
\begin{proposition}
\label{thm:error_estimate}
Consider the inverse problem described by the constrained optimization problem \Cref{eq:invOpt}. Let the form $a:U \times U \to \Reals$ be three times continuously differentiable with respect to the state $u$ and parameters $q$. Let the observation operator $C:U\to\Reals^{n_d}$ be three times continuously differentiable with respect to the state $u$. Also, let the regularization operator $R:Q\to\Reals$ be differentiable with respect to the parameter $q$, and the functional $I:Q\times U\to\Reals$ be differentiable with respect to the state $u$ and parameter $q$.
